\pkg{cubble}: An R Package for Structuring Spatio-temporal Data

### 2022-04-17 2-the-cubble-pkg.Rmd
update codes with the latest cubble

```{r echo = TRUE}
cubble_long <- cubble_nested %>% stretch()
cubble_long
```

@@

```{r echo = TRUE}
cubble_back <- cubble_long %>% tamp()
cubble_back
```

@@

Sometimes, analysts may need to apply some variable transformation that involves both the spatial and temporal variable. An example of this is the transformation of temporal variables into the spatial dimension in glyph maps (which will be elaborated in section \ref{st_transformation}). Cubble allows this operation through \code{migrate()}, which moves the specified spatial variables into the long cubble:

```{r echo = TRUE}
cubble_mig <- cubble_long %>% migrate(long, lat)
cubble_mig
```

@@

```{r echo = TRUE}
cubble_mig %>% tamp() %>% stretch()
```

>>>>

```{r echo = TRUE}
cubble_long <- cubble_nested %>% face_temporal()
cubble_long
```

@@

```{r echo = TRUE}
cubble_back <- cubble_long %>% face_spatial()
cubble_back
```

@@

Sometimes, analysts may need to apply some variable transformation that involves both the spatial and temporal variable. An example of this is the transformation of temporal variables into the spatial dimension in glyph maps (which will be elaborated in section \ref{st_transformation}). Cubble allows this operation through \code{unfold()}, which moves the specified spatial variables into the long cubble:

```{r echo = TRUE}
cubble_mig <- cubble_long %>% unfold(long, lat)
cubble_mig
```

@@

```{r echo = TRUE}
cubble_mig %>% face_spatial() %>% face_temporal()
```

<!-- ## Why not just use dplyrnest_by -->

<!-- ```{r eval = FALSE} -->
<!-- library(cubble) -->
<!-- library(tidyverse) -->
<!-- left_join(cubble::stations, cubble::climate) %>%  -->
<!--   dplyr::nest_by(id, lat,  long,  elev, name, wmo_id) %>%  -->
<!--   tidyr::unnest() %>%  -->
<!--   group_by(........, geometry) -->
<!-- ``` -->

<!-- ```{r eval = FALSE} -->
<!-- as_cubble(list(spatial = stations, temporal = climate), -->
<!--           key = id, index = date, coords = c(long, lat)) -->
<!-- ``` -->

<!-- ```{r eval = FALSE} -->
<!-- usmap::us_map("states") -->
<!-- ``` -->

### 2022-04-17 4-1-basic.Rmd
flesh out section 2.1 - 2.4
update codes with the latest cubble

# Examples

@@

```{r eval=  FALSE, echo = TRUE}
tmax <- tmax %>% 
  stretch() %>%
  filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) 
```

@@

A few stations don't have records during 1971 - 1975 and further investigation shows that while the first and last year of each series is recorded, the missing years in this period is not reported. These stations are filtered out by examining whether the summarised time series has a total of 24 months. The long cubble needs to be switched to the nested form for this spatial operation using \code{tamp()}:

```{r eval = FALSE, echo = TRUE}
tmax <- tmax %>% tamp() %>% filter(nrow(ts) == 24) 
```

@@

```{r eval = FALSE, echo = TRUE}  
tmax <- tmax %>% stretch() %>% migrate(latitude, longitude)
```

>>>>

# Examples {#examples}

@@

```{r eval=  FALSE, echo = TRUE}
tmax <- tmax %>% 
  face_temporal() %>%
  filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) 
```

@@

A few stations don't have records during 1971 - 1975 and further investigation shows that while the first and last year of each series is recorded, the missing years in this period is not reported. These stations are filtered out by examining whether the summarised time series has a total of 24 months. The long cubble needs to be switched to the nested form for this spatial operation using \code{face_spatial()}:

```{r eval = FALSE, echo = TRUE}
tmax <- tmax %>% face_spatial() %>% filter(nrow(ts) == 24) 
```

@@

```{r eval = FALSE, echo = TRUE}  
tmax <- tmax %>% face_temporal() %>% unfold(latitude, longitude)
```

### 2022-04-17 4-3-water-level.Rmd
update codes with the latest cubble

```{r}
climate <- weatherdata::climate_full %>%
  filter(between(stringr::str_sub(id, 7, 8), 76, 90)) %>%
  stretch() %>%
  filter(lubridate::year(date) == 2020) %>%
  tamp() %>%
  mutate(type = "climate")

river <- weatherdata::water %>%
  mutate(type = "river")
```

>>>>

```{r}
climate <- weatherdata::climate_full %>%
  filter(between(stringr::str_sub(id, 7, 8), 76, 90)) %>%
  face_temporal() %>%
  filter(lubridate::year(date) == 2020) %>%
  face_spatial() %>%
  mutate(type = "climate")

river <- river %>% mutate(type = "river")
```

### 2022-04-17 4-5-interactive.Rmd
update codes with the latest cubble

# created SharedData instance for crosstalk
nested <- clean %>% SharedData$new(~id, group = "cubble")
long <- stretch(clean) %>% SharedData$new(~id, group = "cubble")

@@

# create nested form leaflet map with temperature band as subplots 
nested <- tamp(clean)
leaflet(nested) %>% 
  addTiles() %>% 
  addCircleMarkers(group = "a", ...) %>% 
  leafpop::addPopupGraphs(graph = p, ...)

>>>>

# created SharedData instance for crosstalk
nested <- clean %>% SharedData$new(~id, group = "cubble")
long <- face_temporal(clean) %>% SharedData$new(~id, group = "cubble")

@@

# create nested form leaflet map with temperature band as subplots 
nested <- face_spatial(clean)
leaflet(nested) %>% 
  addTiles() %>% 
  addCircleMarkers(group = "a", ...) %>% 
  leafpop::addPopupGraphs(graph = p, ...)

### 2022-04-17 main.tex
small changes
update codes with the latest cubble

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_long <- cubble_nested %>% stretch()
R> cubble_long
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
  id          date        prcp  tmax  tmin
  <chr>       <date>     <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3
2 ASN00009021 2020-01-02     0  24.9  16.4
3 ASN00009021 2020-01-03     6  23.2  13  
4 ASN00009021 2020-01-04     0  28.4  12.4
5 ASN00009021 2020-01-05     0  35.3  11.6
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_back <- cubble_long %>% tamp()
R> cubble_back
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts                
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 x 4]>
2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 x 4]>
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 x 4]>
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 x 4]>
5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 x 4]>
\end{CodeOutput}
\end{CodeChunk}

@@

Sometimes, analysts may need to apply some variable transformation that
involves both the spatial and temporal variable. An example of this is
the transformation of temporal variables into the spatial dimension in
glyph maps (which will be elaborated in section
\ref{st_transformation}). Cubble allows this operation through
\code{migrate()}, which moves the specified spatial variables into the
long cubble:

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_mig <- cubble_long %>% migrate(long, lat)
R> cubble_mig
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
  id          date        prcp  tmax  tmin  long   lat
  <chr>       <date>     <dbl> <dbl> <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3  116. -31.9
2 ASN00009021 2020-01-02     0  24.9  16.4  116. -31.9
3 ASN00009021 2020-01-03     6  23.2  13    116. -31.9
4 ASN00009021 2020-01-04     0  28.4  12.4  116. -31.9
5 ASN00009021 2020-01-05     0  35.3  11.6  116. -31.9
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_mig %>% tamp() %>% stretch()
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
  id          date        prcp  tmax  tmin
  <chr>       <date>     <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3
2 ASN00009021 2020-01-02     0  24.9  16.4
3 ASN00009021 2020-01-03     6  23.2  13  
4 ASN00009021 2020-01-04     0  28.4  12.4
5 ASN00009021 2020-01-05     0  35.3  11.6
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> tmax <- tmax %>% 
+   stretch() %>%
+   filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) 
\end{CodeInput}
\end{CodeChunk}

@@

A few stations don't have records during 1971 - 1975 and further
investigation shows that while the first and last year of each series is
recorded, the missing years in this period is not reported. These
stations are filtered out by examining whether the summarised time
series has a total of 24 months. The long cubble needs to be switched to
the nested form for this spatial operation using \code{tamp()}:

\begin{CodeChunk}
\begin{CodeInput}
R> tmax <- tmax %>% tamp() %>% filter(nrow(ts) == 24) 
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> tmax <- tmax %>% stretch() %>% migrate(latitude, longitude)
\end{CodeInput}
\end{CodeChunk}

@@

\hypertarget{interative-graphic-with-cubble}{%
\subsection{Interative graphic with
cubble}\label{interative-graphic-with-cubble}}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> # data pre-processing
R> clean <- weatherdata::climate_full %>% ...
R> 
R> # created SharedData instance for crosstalk
R> nested <- clean %>% SharedData$new(~id, group = "cubble")
R> long <- stretch(clean) %>% SharedData$new(~id, group = "cubble")
R> 
R> # create the spatial and temporal view each with a ShareData instance
R> p1 <- nested %>% ...
R> p2 <- long %>% ...
R> 
R> # Combine p1 and p2
R> crosstalk::bscols(plotly::ggplotly(p1), plotly::ggplotly(p2), ...)
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> # data pre-processing
R> clean <- weatherdata::climate_full %>% ...
R> 
R> # use the long form to create subplots for each station
R> df_id <- unique(clean$id)
R> p <- map(1:length(df_id), function(i){
+   dt <- clean %>% filter(id == df_id[i])
+   ggplot(dt) %>% ...
+ })
R> 
R> # create nested form leaflet map with temperature band as subplots 
R> nested <- tamp(clean)
R> leaflet(nested) %>% 
+   addTiles() %>% 
+   addCircleMarkers(group = "a", ...) %>% 
+   leafpop::addPopupGraphs(graph = p, ...)
\end{CodeInput}
\end{CodeChunk}

>>>>

%% recommended packages
\usepackage{orcidlink,thumbpdf,lmodern}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_long <- cubble_nested %>% face_temporal()
R> cubble_long
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
  id          date        prcp  tmax  tmin
  <chr>       <date>     <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3
2 ASN00009021 2020-01-02     0  24.9  16.4
3 ASN00009021 2020-01-03     6  23.2  13  
4 ASN00009021 2020-01-04     0  28.4  12.4
5 ASN00009021 2020-01-05     0  35.3  11.6
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_back <- cubble_long %>% face_spatial()
R> cubble_back
\end{CodeInput}
\begin{CodeOutput}
# cubble:   id [5]: nested form
# bbox:     [115.97, -32.94, 133.55, -12.42]
# temporal: date [date], prcp [dbl], tmax [dbl], tmin [dbl]
  id            lat  long  elev name           wmo_id ts                
  <chr>       <dbl> <dbl> <dbl> <chr>           <dbl> <list>            
1 ASN00009021 -31.9  116.  15.4 perth airport   94610 <tibble [366 x 4]>
2 ASN00010311 -31.9  117. 179   york            94623 <tibble [366 x 4]>
3 ASN00010614 -32.9  117. 338   narrogin        94627 <tibble [366 x 4]>
4 ASN00014015 -12.4  131.  30.4 darwin airport  94120 <tibble [366 x 4]>
5 ASN00015131 -17.6  134. 220   elliott         94236 <tibble [366 x 4]>
\end{CodeOutput}
\end{CodeChunk}

@@

Sometimes, analysts may need to apply some variable transformation that
involves both the spatial and temporal variable. An example of this is
the transformation of temporal variables into the spatial dimension in
glyph maps (which will be elaborated in section
\ref{st_transformation}). Cubble allows this operation through
\code{unfold()}, which moves the specified spatial variables into the
long cubble:

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_mig <- cubble_long %>% unfold(long, lat)
R> cubble_mig
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
  id          date        prcp  tmax  tmin  long   lat
  <chr>       <date>     <dbl> <dbl> <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3  116. -31.9
2 ASN00009021 2020-01-02     0  24.9  16.4  116. -31.9
3 ASN00009021 2020-01-03     6  23.2  13    116. -31.9
4 ASN00009021 2020-01-04     0  28.4  12.4  116. -31.9
5 ASN00009021 2020-01-05     0  35.3  11.6  116. -31.9
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> cubble_mig %>% face_spatial() %>% face_temporal()
\end{CodeInput}
\begin{CodeOutput}
# cubble:  date, id [5]: long form
# bbox:    [115.97, -32.94, 133.55, -12.42]
# spatial: lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]
  id          date        prcp  tmax  tmin
  <chr>       <date>     <dbl> <dbl> <dbl>
1 ASN00009021 2020-01-01     0  31.9  15.3
2 ASN00009021 2020-01-02     0  24.9  16.4
3 ASN00009021 2020-01-03     6  23.2  13  
4 ASN00009021 2020-01-04     0  28.4  12.4
5 ASN00009021 2020-01-05     0  35.3  11.6
# ... with 1,825 more rows
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> tmax <- tmax %>% 
+   face_temporal() %>%
+   filter(lubridate::year(date) %in% c(1971:1975, 2016:2020)) 
\end{CodeInput}
\end{CodeChunk}

@@

A few stations don't have records during 1971 - 1975 and further
investigation shows that while the first and last year of each series is
recorded, the missing years in this period is not reported. These
stations are filtered out by examining whether the summarised time
series has a total of 24 months. The long cubble needs to be switched to
the nested form for this spatial operation using \code{face_spatial()}:

\begin{CodeChunk}
\begin{CodeInput}
R> tmax <- tmax %>% face_spatial() %>% filter(nrow(ts) == 24) 
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> tmax <- tmax %>% face_temporal() %>% unfold(latitude, longitude)
\end{CodeInput}
\end{CodeChunk}

@@

\hypertarget{interative-graphic}{%
\subsection{Interative graphic}\label{interative-graphic}}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> # data pre-processing
R> clean <- weatherdata::climate_full %>% ...
R> 
R> # created SharedData instance for crosstalk
R> nested <- clean %>% SharedData$new(~id, group = "cubble")
R> long <- face_temporal(clean) %>% SharedData$new(~id, group = "cubble")
R> 
R> # create the spatial and temporal view each with a ShareData instance
R> p1 <- nested %>% ...
R> p2 <- long %>% ...
R> 
R> # Combine p1 and p2
R> crosstalk::bscols(plotly::ggplotly(p1), plotly::ggplotly(p2), ...)
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> # data pre-processing
R> clean <- weatherdata::climate_full %>% ...
R> 
R> # use the long form to create subplots for each station
R> df_id <- unique(clean$id)
R> p <- map(1:length(df_id), function(i){
+   dt <- clean %>% filter(id == df_id[i])
+   ggplot(dt) %>% ...
+ })
R> 
R> # create nested form leaflet map with temperature band as subplots 
R> nested <- face_spatial(clean)
R> leaflet(nested) %>% 
+   addTiles() %>% 
+   addCircleMarkers(group = "a", ...) %>% 
+   leafpop::addPopupGraphs(graph = p, ...)
\end{CodeInput}
\end{CodeChunk}
